---
title: "Problem Set 1"
author: "Aastha"
date: "1/27/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r }
library(tidyverse)
library(readr)
library(ks)
library(kdensity)
library(data.table)
library(fitdistrplus)
library(ggplot2)
library(gridExtra)
library(ggpubr)

setwd("~/github/empirical-methods/ProblemSets/pset1")

dt <- read_csv("bids1.csv", col_names = c("bids"))
```
# Question 1

### Histogram of the bids
```{r}
ggplot (data = dt, aes(dt$bids)) + geom_histogram(binwidth = 0.25)



```


```{r , echo=FALSE}
qplot(dt$bids, geom="histogram", binwidth = 0.25,main = "Histogram of bid values", xlab = "Bid values", ylab = "Frequency",  position = "center", col=I("red"), fill = I("white"))

```
# Question 2

1. Estimating bids assuming normal distribution:

```{r}
fit_normal <- fitdist(dt$bids, "norm")
class(fit_normal)
#plot(fit_normal)
denscomp(fit_normal, xlab = "Bid values")

```
Alternate way for normal density:

```{r}
mean(dt$bids)# 2.19
sd(dt$bids)# 0.74

density_normal = dnorm(dt$bids, mean = mean(dt$bids), sd =sd(dt$bids))
plot(dt$bids, density_normal, xlab = "Bid values" , ylab = "density")

```

2. Estimating bids using a Guassian Kernel


```{r}

density_guassian <- density(dt$bids, bw = "nrd0", adjust = 1,
        kernel = "gaussian",
        weights = NULL, window = kernel, width,
        give.Rkern = FALSE,
        n = 512,  cut = 3, na.rm = FALSE)

plot(density_guassian, xlab="Bid values", main = "Gaussian Kernel")

```



3. Estimating bids uding Epanechnikov Kernel


```{r}
density_epanechnikov <- density(dt$bids, bw = "nrd0", adjust = 1,
                                kernel = "epanechnikov",
                                weights = NULL, window = kernel, width,
                                give.Rkern = FALSE,
                                n = 512,  cut = 3, na.rm = FALSE)



plot(density_epanechnikov, xlab="Bid values", main = "Epanechnikov Kernel")


```


# Question 2 

Using a least-squares cross-validation to pick the bandwidth for the Epanechnikov kernel


```{r}
bids <- dt$bids
X = bids

# optimal bandwidth

IQR(X)
sd(X)
opt_h = 0.9*sd(X)*length(X)^(-1/5) # optimal h


J=function(h){
   fhat=Vectorize(function(x) density(X,from=x,to=x,n=1,bw=h)$y)
   fhati=Vectorize(function(i) density(X[-i],from=X[i],to=X[i],n=1,bw=h)$y)
   F=fhati(1:length(X))
   return(integrate(function(x) fhat(x)^2,-Inf,Inf)$value-2*mean(F))
   }

vx=seq(0.1,1,by=.01)
vy=Vectorize(J)(vx)
df=data.frame(vx,vy)
qplot(vx,vy,geom="line",data=df)

optimize(J,interval=c(.1,1)) 


````
We obtain optimal h = 0.317 by least squares cross validation, using which we obtain the following density function.


```{r}
density_epanechnikov_CV <- density(dt$bids, bw=0.31, adjust = 1, 
                   kernel = c("epanechnikov"),
                   weights = NULL)

plot(density_epanechnikov_CV,xlab="Bid values", main = "Epanechnikov kernel after cross validation" )


````

# Question 3
Trying to find the best fit of density for the data
```{r}

density_normal1 <- rnorm(length(dt$bids), mean(dt$bids), sd(dt$bids))

hist(dt$bids, breaks = seq(0.5,4,0.25), freq = FALSE, main = "Comparing different density functions", xlab = "Bid values")
lines(density(density_normal1, adjust = 2), col ="chocolate", lwd =2) ### some problem here
lines(density_epanechnikov, col="red", lwd = 2)
lines(density_epanechnikov_CV, col ="green", lwd =2)
lines(density_guassian, col = "blue", lwd =2)

legend(1, 95, legend=c("normal", "epanechnikov", "epanechnikov_CV", "gaussian"),
       col=c("chocolate", "red", "green", "blue"), lty=1:2, cex=0.8)

```
chocolate = normal distribution, red = epanechnikov, green = epnechnikov after cross validation, blue = gaussian

# Question 4

```{r}

class(density_epanechnikov_CV)
max(density_epanechnikov_CV$y)
sum(density_epanechnikov_CV$y)
#density_epanechnikov_CV = density_epanechnikov_CV$y/sum(density_epanechnikov_CV$y)

library(spatstat)
#density_epanechnikov_CV <- as.function(density_epanechnikov_CV)
cdf = CDF(density_epanechnikov_CV)
plot(cdf)

```

```{r}

# calculating cdf and pdf g_hat and G_hat
g_hat <- approxfun(density_epanechnikov_CV)
yGb <- c(cumsum(density_epanechnikov_CV$y))
xGb <- c(density_epanechnikov_CV$x)
Gb  <- data.frame(yGb,xGb)
G_hat <- approxfun(Gb)


v_hat <- function(b) {
  results <-    b + (1/(2))*(G_hat(b)/g_hat(b)) # using the GPV formula
  return(results)
}
vHat <- v_hat(dt$bids)
min(vHat)
max(vHat)



```

# Question 5

Estimating the values "v" using epanechnikov kernel


```{r}
#hist(vHat, breaks = seq(2,10,0.5), freq = FALSE)
density_vHat = density(vHat, bw="nrd0", adjust = 1, 
                       kernel = c("epanechnikov"),
                       weights = NULL)


hist(vHat, breaks = seq(2,10,0.5), freq = FALSE, main = "Distribution of bid values")
lines(density_vHat, col = "red")


```


#Question 6


```{r}
fitg <- fitdist(vHat, "gamma") # loglikelihood -437
fitln <- fitdist(vHat, "lnorm")
fitw <- fitdist(vHat, "weibull")


summary(fitg)


hist(vHat, breaks = seq(2,10,0.5), freq = FALSE, main = "Distribution of bid values")
lines(density_vHat, col = "red")



```

```{r}
denscomp(fitg)
lines(density_vHat, col = "blue")

denscomp(fitln)
lines(density_vHat, col = "purple")

denscomp(fitw)
lines(density_vHat, col = "green")


```


```{r}

summary(fitg)
summary(fitln)
summary(fitw)

```

The distribution closely represents the lognormal distribution as it appears from the above comparsion of density graphs. 